Stage-Splitting a Modal Language

### 2014-06-05 paper/paper.tex
some more text on semantics

In this section, we introduce a dynamic semantics for our language.  Although our type system is virtually identical to that in \cite{davies96}, the semantics is different in terms of cost and termination behavior.  The semantics comprises the following judgments:

@@

Evaluation and speculation are the main judgments here, the rest being largely administrative.  The evaluation judgment operates on stage-1 code whereas the speculation operates on stage-2 code.  Since $\next$ and $\prev$ are the crossover points between 1-code and 2-code, they are correspondingly the only places where the evaluation and speculation judgments depend on the other.

We can think of our evaluation as proceeding in the standard way for stage-1 code. But when the evaluator encounters stage-2 code, it places that code off to the side in a table, and then keeps a reference to the table entry.  These references are then passed around as stage-1 values, and the tables are managed accordingly.

The evaluation judgment is very similar to standard call-by-value evaluation, as goal 1 would suggest.  The input to evaluation is a stage-1 expression (usually $e$), as well as two administrative contexts ($\Gamma$ and $\Xi$), covered later.  Evaluation has two outputs: the {\em future table} (usually $\xi$) and the {\em partial value} (usually $v$).  We cover the latter first.  The partial value is essentially the result of the first-stage portion of $e$, and must be a 1-value of the same type as $e$.  As usual, this means that $v$ must be composed only of base primitives, tuples, injections, and lambdas (corresponding to base types, products, sums, and functions).  Analogously, the value corresponding to $\fut$ is a construct called a {\em hatted variable} (written $\hat y$), which signifies a reference to some stage-2 computation.  Those stage-2 computations are held in

@@

>>>>

In this section, we introduce a dynamic semantics for our language.  Although our type system is virtually identical to that in \cite{davies96}, the semantics is different in terms of cost and termination behavior.

Abstractly, we can think of our evaluation as proceeding in the standard way for stage-1 code. When the evaluator encounters a $\next \{e\}$ expression, it places $e$ off to the side in a table and replaces the whole expression with a reference to the table entry.  These references are then passed around as stage-1 values for $\fut$ types.  But what if $e$ contains $\prev$ expressions?  To ensure that all stage-1 code is evaluated before any stage-2 code, we must evaluate all of the 1-code contained in $e$ before inserting it into the table.  This entails searching $e$ for all contained $\prev$s and evaluating them in place.

More formally, evaluation comprises the folowing judgments:

@@

Evaluation and speculation are the main judgments here, the rest being largely administrative.  The evaluation judgment operates on stage-1 code, whereas the speculation operates on stage-2 code.  Since $\next$ and $\prev$ are the crossover points between 1-code and 2-code, they are correspondingly the only places where the evaluation and speculation judgments depend on the other.

The evaluation judgment is very similar to standard call-by-value evaluation, as goal 1 would suggest.  The input to evaluation is a stage-1 expression (usually $e$), as well as two administrative contexts ($\Gamma$ and $\Xi$), covered later.  Evaluation has two outputs: the {\em future table} (usually $\xi$) and the {\em partial value} (usually $v$).  We cover the latter first.  The partial value is essentially the result of the first-stage portion of $e$, and must be a 1-value of the same type as $e$.  As usual, this means that $v$ must be composed only of base primitives, tuples, injections, and lambdas (corresponding to base types, products, sums, and functions).  Analogously, the value corresponding to $\fut$ is a construct called a {\em hatted variable} (written $\hat y$), which signifies a reference to some stage-2 computation.

...

@@

\subsection{Comparison to Other Semantics}

The differences between our diagonal semantics, the erasure semantics, and Davies' semantics can be characterized entirely by how they behave at $\next$ expressions:
\begin{itemize}
\item {\bf Erasure Semantics.} Evaluate contents of $\next$ immediately.
\item {\bf Diagonal Semantics.} Speculate to find the $\prev$s, evaluate them in place, then place the contents of $\next$ into a table and create a hatted variable to act as a value.
\item {\bf Davies' Semantics.} Speculate to find the $\prev$s, evaluate them in place, then treat the reduced $\next$ as a value.
\end{itemize}
